\newif\ifarxiv
\arxivtrue % swap this to true to build for the arxiv
\ifarxiv
\documentclass{article}
\usepackage{arxiv}
\else
\documentclass[review,onefignum,onetabnum]{siamonline220329}
\fi

\pdfoutput=1
% below is for arxiv
% \documentclass{article}
% \pdfoutput=1
% \usepackage{arxiv}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\ifarxiv
\else
\ifpdf
\hypersetup{
  pdftitle={Let them have CAKES: A Cutting-Edge Algorithm for Scalable, Efficient, and Exact Search on Big Data},
  pdfauthor={M. E. Prior, T. J. Howard III, O. McLaughlin, T. Ferguson, N. Ishaq, N. M. Daniels}
}
\fi
\fi





\input{cakes_shared}
\begin{document}


\maketitle



\begin{abstract}
  The Big Data explosion has created a demand for efficient and scalable algorithms for similarity search.
  While much recent work has focused on \textit{approximate} $k$-NN search, \textit{exact} $k$-NN search has not kept up.
  We present CAKES, a set of three novel algorithms for exact $k$-NN search.
  CAKES's algorithms are generic over \textit{any} distance function, and do not scale with the cardinality or embedding dimension of the dataset. Instead, they scale with geometric properties of the dataset--namely, metric entropy and fractal dimension--thus providing immense speed improvements over existing exact $k$-NN search algorithms when the dataset conforms to the manifold hypothesis.
  We demonstrate these claims by contrasting performance on a randomly-generated dataset against that on some datasets from the ANN-Benchmarks suite under commonly-used distance functions, a genomic dataset under Levenshtein distance, and a radio-frequency dataset under Dynamic Time Warping distance. CAKES exhibits near-constant running time on data conforming to the manifold hypothesis as cardinality grows, and has perfect recall on data in metric spaces. CAKES also has significantly higher recall than state-of-the-art $k$-NN search algorithms even when the distance function is not a metric.
  We conclude that CAKES is a highly efficient and scalable algorithm for exact $k$-NN search on Big Data.
  We provide a Rust implementation of CAKES under an MIT license at https://github.com/URI-ABD/clam.

\end{abstract}
%
% \keywords{K-NN Search, Manifold Hypothesis, Sub-Linear  Algorithms, Big Data}

% REQUIRED
\ifarxiv
\else
\begin{MSCcodes}
68P05, 68P10
\end{MSCcodes}
\fi
%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

    \input{sections/1-introduction.tex}
    \input{sections/2-methods.tex}
    \input{sections/3-datasets.tex}
    \input{sections/4-results.tex}
    \input{sections/5-discussion.tex}

    %\section*{Acknowledgments}
    %The authors thank the members of the University of Rhode Island's Algorithms for Big Data research group for their helpful comments throughout the development of this work.
    %We are especially grateful to Carl Stoker and Rachel F. Daniels for their thorough reviews of the paper and valuable feedback.

    % \afterpage{\clearpage}
    \FloatBarrier
    \bibliographystyle{siamplain}
    \bibliography{references}
    % \newpage
    % \input{supplement.tex}
\end{document}
